FM project

### 2023-05-10 draft1.tex
updeited

---
title: "FM project"
output:
  html_document:
    citation_package: citr
date: "2023-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

@@

https://www.eia.gov/todayinenergy/detail.php?id=55920#:~:text=Europe%20became%20the%20primary%20destination,U.S.%20LNG%20exports%20to%20Europe.

## Analysing U.S. energy exports

### 1.Introduction

@@

### 2. Time Series Graphics

@@

#### 2.1 Dataset

The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.

@@

```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
library(fpp3)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(scales)

#Total Crude Oil Exports by Destination

raw_crude_oil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Crude%20Oil%20Exports%20by%20Destination_EIA.csv", skip = 2,)

#Leaving only country names in column names

names(raw_crude_oil)[-2] <- str_replace_all(names(raw_crude_oil)[-2],
                                            c("U.S. Exports to " = "",
                                              " of Crude Oil \\(Thousand Barrels\\)" = ""))

#Pivoting (switching) columns starting from 2 column to rows

crude_oil <- raw_crude_oil %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of Crude Oil (Thousand Barrels)')

#Filtering out values where Date is 'NA'

crude_oil <- crude_oil %>%
  drop_na(Date) %>%
  mutate(Date = as.Date(Date))

#Fixing date format, replacing 'na' values with 0

crude_oil$Date <- as.Date(crude_oil$Date)
crude_oil$`Amount of Crude Oil (Thousand Barrels)` <- replace_na(crude_oil$`Amount of Crude Oil (Thousand Barrels)`, 0)

#Setting up the tsibble

ts_crude_oil <- crude_oil %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date, 
    key = `Destination country`
  )

#OECD Europe Countries vector

oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
                 "United Kingdom")

#filtering out only relevant values

oecd_crude_oil <- ts_crude_oil %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

oecd_crude_oil_agg <- oecd_crude_oil %>%
  index_by(Date) %>%
  summarize_at(vars(`Amount of Crude Oil (Thousand Barrels)`), sum) %>%
  ungroup()

################################################################################

#Total Oil Products Exports by Destination

raw_oil_prod <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Oil%20Products%20Exports%20by%20Destination_EIA.csv",  skip = 2,)

names(raw_oil_prod)[-2] <- str_replace_all(names(raw_oil_prod)[-2],
                                           c("U.S. Exports to " = "",
                                             " of Total Petroleum Products \\(Thousand Barrels\\)" = ""))
oil_prod <- raw_oil_prod %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of total Petroleum Porducts (Thousand Barrels)')

oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)` <- replace_na(oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0)

oil_prod <- oil_prod %>% 
  drop_na(Date) %>% 
  mutate(Date = as.Date(Date))

ts_oil_prod <- oil_prod %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date, 
    key = `Destination country`
  )

oecd_oil_prod <- ts_oil_prod %>% 
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

oecd_oil_prod_agg <- oecd_oil_prod %>%
  index_by(Date) %>%
  summarize_at(vars(`Amount of total Petroleum Porducts (Thousand Barrels)`), sum) %>%
  ungroup()

################################################################################
#Reliance on Russian gas

raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)

rel_rusoil <- raw_rel_rusoil %>% 
  pivot_longer(cols = -...1,
               names_to = 'Date',
               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')

rel_rusoil <- rel_rusoil %>% 
  rename(Country = ...1) %>% 
  mutate(Date = lubridate::dmy(Date)) %>% 
  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")

ts_rel_rusoil <- rel_rusoil %>% 
  as_tsibble(
    index = Date,
    key = Country)

oecd_relrus <- ts_rel_rusoil %>%
  filter(Country %in% oecd_europe)

################################################################################
#Total Natural Gas Exports by Destination

raw_nat_gas <- read.csv('https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Natural%20Gas%20Exports%20by%20Country_EIA.csv',header=TRUE , check.names=FALSE, skip = 2)

# drop columns 2-3, 6-8, 54 and 57-72
raw_nat_gas <- subset(raw_nat_gas, select = -c(2:3, 6:8,54, 57:72))
raw_nat_gas[is.na(raw_nat_gas)] <- 0

#cimbining exports to the same countries

nat_gas <- raw_nat_gas %>%
  mutate(Mexico = `U.S. Natural Gas Pipeline Exports to Mexico (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)`,
         Canada = `U.S. Natural Gas Pipeline Exports to Canada (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)`)

#removing duplicate columns

nat_gas <- nat_gas[, !(names(nat_gas) %in% c('U.S. Natural Gas Pipeline Exports to Mexico (MMcf)',
                                             'Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)',
                                             'U.S. Natural Gas Pipeline Exports to Canada (MMcf)',
                                             'Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)'))]

#cleaning column names

colnames(nat_gas) <- gsub(pattern = ".*(to\\s)", "", colnames(nat_gas))
colnames(nat_gas) <- gsub("\\(.*?\\)", "", colnames(nat_gas))

#pivoting data
gastemp <- nat_gas %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of Natural gas (MMcf)')

#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)
library(stringr) #for triming the strings

gastemp <- gastemp %>% 
  drop_na(Date) %>% 
  mutate(Date = as.Date(Date)) %>% 
  mutate(`Destination country` = str_trim(`Destination country`))

#summing duplicates

gastemp <- gastemp %>%
  group_by(Date, `Destination country`) %>%
  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
  ungroup() %>% 
  drop_na(Date)

#creating a tsibble

ts_nat_gas <- gastemp %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date,
    key = `Destination country`
  )

#filtering only relevant data

oecd_nat_gas <- ts_nat_gas %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2010-01-15"), Date <= as.Date("2023-01-15"))

oecd_nat_gas_agg <- oecd_nat_gas %>%
  index_by(Date) %>%
  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
  ungroup()

oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
  mutate(`Amount of Natural gas (BOE)` = `Amount of Natural gas (MMcf)` * 1000 / 6000 )

oecd_nat_gas_agg <- oecd_nat_gas_agg %>% 
  select(Date, `Amount of Natural gas (BOE)`)

################################################################################

oecd_merged <- left_join(oecd_nat_gas_agg, oecd_oil_prod_agg, by = "Date") %>%
  left_join(oecd_crude_oil_agg, by = "Date")

total_exp_oecd <- oecd_merged %>%
  pivot_longer(cols = -Date, names_to = "Export Type", values_to = "Amount")

total_exp_oecd %>% ggplot(
 aes(Date,Amount,color = `Export Type`)) +
   geom_line()+
   labs(x = "Date", y = "Amount")+
  theme(legend.position = "bottom")

```
**Comment on Nat Gas export zeros**

#### 2.2 Time Series Plots

@@

#### Including Plots

@@

```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE}

total_exp_oecd %>% ggplot(
 aes(Date,Amount,color = `Export Type`)) +
   geom_line()+
   labs(x = "Date", y = "Amount")+
  theme(legend.position = "bottom")

```

@@

```{r, overview by product, echo=TRUE, results='markup',warning=FALSE,message=FALSE}

total_exp_oecd %>%  
  ggplot(aes(Date, Amount, color = `Export Type`)) +
  geom_line()+
  labs(x = "Date", y = "Amount")+
  theme(legend.position = "bottom") +
  facet_wrap(~`Export Type`, ncol = 1)

```
#### 2.3 Time Series Patterns
```{r, Seasonal plot for each product, echo=FALSE, results='markup',warning=FALSE,message=FALSE}
# create seasonal plot of total exports monthly of Natural gas
total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_season(Total_Exports) +
  ylab("Total exports  of Natural Gas") +
  ggtitle("Seasonal plot: US exports of natural gas to OECD") +
  scale_y_continuous(labels = comma_format())

# create seasonal plot of total exports monthly of Crude oil
total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_season(Total_Exports) +
  ylab("Total exports  of Crude oil") +
  ggtitle("Seasonal plot: US exports of Crude Oil to OECD") +
  scale_y_continuous(labels = comma_format())

# create seasonal plot of total exports monthly of Petroleum products
total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_season(Total_Exports) +
  ylab("Total exports  of Petroleum Products") +
  ggtitle("Seasonal plot: US exports of Petroleum to OECD") +
  scale_y_continuous(labels = comma_format())
```
To further elaborate, the data analysis of US exports of oil products to OECD shows that seasonal patterns can differ depending on the type of product being exported. While natural gas and crude oil both display some level of seasonality, petroleum products do not show any significant trends.

In the case of natural gas, we can see a clear increase in exports during the colder months, which could be due to increased demand for heating purposes. Conversely, during the warmer months, exports tend to decrease. This pattern is consistent throughout the years, with the exception of some yearly fluctuations.

Regarding crude oil, the seasonal pattern is not as clear as with natural gas. However, we still observe a similar trend where exports tend to increase during the colder months and decrease during the warmer months. Interestingly, we see a deviation from this pattern in 2022, with exports remaining high even during spring and summer months. This could be attributed to the political tensions caused by the Russian invasion in Ukraine.

On the other hand, the export of petroleum products does not appear to have any notable seasonal patterns. While there may be a slight increase towards the end of each year, there is no clear trend observed year-over-year.

Overall, analyzing seasonal patterns in the US export of oil products can provide insights into demand patterns and other underlying factors that affect the global oil market.

```{r, seasonal subsidaries plots,box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE }
#Subseries plots Natural gas
total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_subseries(Total_Exports) + 
  ylab("Total exports") + 
  ggtitle("Subseries plot: US exports of natural gas")+
  scale_y_continuous(labels = comma_format())

#Subseries plots Crude oil
total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_subseries(Total_Exports) + 
  ylab("Total exports") + 
  ggtitle("Subseries plot: US exports of Crude Oil")+
  scale_y_continuous(labels = comma_format())

#Subseries plots Petroleum products
total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  gg_subseries(Total_Exports) + 
  ylab("Total exports") + 
  ggtitle("Subseries plot: US exports of Petroleum Products")+
  scale_y_continuous(labels = comma_format())
```
To gain a better understanding of seasonality, we created additional plots known as subseries seasonal plots. These graphs provide a clearer view of monthly exports and also display the means over each month for the analyzed time series period.

When we look at the subseries seasonal plot for natural gas exports, we can see a clear seasonality of increased natural gas exports during the colder seasons, starting from September and decreasing in spring. From the crude oil subseries seasonal plot, we can notice a huge increase in December and January as well, indicating that these months have a strong seasonal effect on crude oil exports. Finally, from the subseries seasonal plot of Petroleum products, we can notice the same increase in December, but there is a huge decrease in February and March. All other months are fluctuating more randomly without displaying any clear seasonality.

In general, the seasonal subseries plots for all three types of oil exports to OECD reveal similar patterns to those seen in the seasonal plots. There is a clear increase in exports during the colder seasons, particularly in December and January, and a decrease during the warmer seasons. However, there are some fluctuations and irregularities throughout the year that do not follow a clear seasonal pattern.

####2.4 ACF and White Noise

```{r, }
#ACF plot without lag Natural Gas

total_exp_oecd %>% filter(`Export Type` == "Amount of Natural gas (BOE)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  ACF() %>%  autoplot()

#ACF plot without lag Crude Oil

total_exp_oecd %>% filter(`Export Type` == "Amount of Crude Oil (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  ACF() %>%  autoplot()

#ACF plot without lag Petroleum

total_exp_oecd %>% filter(`Export Type` == "Amount of total Petroleum Porducts (Thousand Barrels)") %>% 
  index_by(Date) %>% 
  summarise(Total_Exports = sum(`Amount`)) %>% 
  ACF() %>%  autoplot()

```

The autocorrelation function (ACF) plot displays the correlation between a time series and its own lagged values. A high ACF value for a particular lag indicates that the current values of the time series are significantly correlated with the past values at that lag.

The ACF plots for US exports of Natural Gas, Crude Oil to OECD reveal interesting insights regarding the seasonality and trends in the time series. In the case of Natural Gas and Crude Oil, the ACF plots demonstrate a significant positive correlation, indicating the presence of high trend component with slight seasonality patterns observed as. However, these plots reflect a dominant upward trend, as evidenced by the substantial increase in exports. The trend component seems to be overtaking the seasonality, suggesting that the overall upward trend in exports is more prominent than the periodic variation in the data.

The ACF plot for US exports of Petroleum Products to OECD indicates that there is no noticeable seasonality in the time series. The first three lags show significant positive correlations, suggesting some short-term dependence between current exports and those from 1 to 3 months ago. However, beyond the third lag, the ACF values fall within the confidence interval, indicating a weakening of the correlation. This suggests that there is no clear and consistent pattern of seasonality in the exports of petroleum products.

Finally, based on the ACF plots of the US exports of natural gas, crude oil, and petroleum products to OECD, we can conclude that the time series do not appear to be white noise.

### 3. Time series decomposition
#### 3.1. Adjustments

@@

#### 3.2. Transformations

@@

```{r, box_cox, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

lambda1 <- oecd_crude_oil_agg %>% 
  features(`Amount of Crude Oil (Thousand Barrels)`, features = guerrero) %>% 
  pull(lambda_guerrero)
crude_oil_agg <- oecd_crude_oil_agg %>% 
  mutate(`Amount of Crude Oil (Thousand Barrels)` = box_cox(`Amount of Crude Oil (Thousand Barrels)`, lambda1))

lambda2 <- oecd_oil_prod_agg %>% 
  features(`Amount of total Petroleum Porducts (Thousand Barrels)`, features = guerrero) %>% 
  pull(lambda_guerrero)
oil_prod_agg <-oecd_oil_prod_agg %>% 
  mutate(`Amount of total Petroleum Porducts (Thousand Barrels)` = box_cox(`Amount of total Petroleum Porducts (Thousand Barrels)`, lambda2))

lambda3 <- oecd_nat_gas_agg %>% 
  features(`Amount of Natural gas (BOE)`, features = guerrero) %>% 
  pull(lambda_guerrero)
nat_gas_agg <- oecd_nat_gas_agg %>% 
  mutate(`Amount of Natural gas (BOE)` = box_cox(`Amount of Natural gas (BOE)`, lambda3))

```

#### 3.3. Decomposition

###STL Decomposition

```{r, decomposition, echo=FALSE, results='markup',warning=FALSE,message=FALSE }

dcmp_oecd_crude_oil <- oecd_crude_oil_agg %>%
  model(stl = STL(`Amount of Crude Oil (Thousand Barrels)`))

components(dcmp_oecd_crude_oil)

dcmp_oecd_oil_prod <- oecd_oil_prod_agg %>%
  model(stl = STL(`Amount of total Petroleum Porducts (Thousand Barrels)`))

components(dcmp_oecd_oil_prod)

dcmp_oecd_nat_gas <- oecd_nat_gas_agg %>%
  model(stl = STL(`Amount of Natural gas (BOE)`))

components(dcmp_oecd_nat_gas)

```

```{r, decomposition_plots, echo=FALSE, results='markup',warning=FALSE,message=FALSE }

components(dcmp_oecd_crude_oil) %>% autoplot()

components(dcmp_oecd_oil_prod) %>% autoplot()

components(dcmp_oecd_nat_gas) %>% autoplot()
```

@@

## 4. The forecaster’s toolbox

#### 4.1. Exploring different simple forecasting methods

Before moving forward with forecasting using simple forecasting methods we take a look at both time series and what simple forecasting methods works best with both

```{r, evaluating models,echo=FALSE, results='markup',warning=FALSE,message=FALSE  }

crude_oil_train <- crude_oil_agg %>% 
  filter(Date >= as.Date("2015-06-01"), Date <= as.Date("2021-06-01"))

#fiting all four models

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={FM project},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{FM project}
\author{}
\date{\vspace{-2.5em}2023-04-25}

\begin{document}
\maketitle

@@

\url{https://www.eia.gov/todayinenergy/detail.php?id=55920\#}:\textasciitilde:text=Europe\%20became\%20the\%20primary\%20destination,U.S.\%20LNG\%20exports\%20to\%20Europe.

\hypertarget{analysing-u.s.-energy-exports}{%
\subsection{Analysing U.S. energy
exports}\label{analysing-u.s.-energy-exports}}

\hypertarget{introduction}{%
\subsubsection{1.Introduction}\label{introduction}}

@@

\hypertarget{time-series-graphics}{%
\subsubsection{2. Time Series Graphics}\label{time-series-graphics}}

@@

\hypertarget{dataset}{%
\paragraph{2.1 Dataset}\label{dataset}}

The data sets utilized in this study consist of monthly observations of
Country's reliance on Russian oil products, USA's natural gas exports by
country, USA's total crude oil exports by destination, and USA's
petroleum products exports by destination.

@@

\includegraphics{draft1_files/figure-latex/data_prep-1.pdf}
\textbf{Comment on Nat Gas export zeros}

\hypertarget{time-series-plots}{%
\paragraph{2.2 Time Series Plots}\label{time-series-plots}}

@@

\hypertarget{including-plots}{%
\paragraph{Including Plots}\label{including-plots}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}
 \FunctionTok{aes}\NormalTok{(Date,Amount,}\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
   \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
   \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft1_files/figure-latex/time series overview-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{total\_exp\_oecd }\SpecialCharTok{\%\textgreater{}\%}  
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Date, Amount, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{()}\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Amount"}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\StringTok{\textasciigrave{}}\AttributeTok{Export Type}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{draft1_files/figure-latex/overview by product-1.pdf}
\#\#\# 3. Time series decomposition \#\#\#\# 3.1. Adjustments

@@

\hypertarget{transformations}{%
\paragraph{3.2. Transformations}\label{transformations}}

@@

\hypertarget{decomposition}{%
\paragraph{3.3. Decomposition}\label{decomposition}}

\#\#\#STL Decomposition

\begin{verbatim}
## # A dable: 156 x 7 [1M]
## # Key:     .model [1]
## # :        Amount of Crude Oil (Thousand Barrels) = trend + season_year +
## #   remainder
##    .model     Date Amount of Crude Oil (Thousan~1  trend seaso~2 remai~3 seaso~4
##    <chr>     <mth>                          <dbl>  <dbl>   <dbl>   <dbl>   <dbl>
##  1 stl    2010 Feb                              0 -20.8    -31.4    52.2    31.4
##  2 stl    2010 Mar                              0 -18.7    222.   -203.   -222. 
##  3 stl    2010 Apr                              0 -16.6     41.9   -25.3   -41.9
##  4 stl    2010 May                              0 -14.4    383.   -369.   -383. 
##  5 stl    2010 Jun                              0 -13.1   -326.    339.    326. 
##  6 stl    2010 Jul                              0 -11.9   -435.    447.    435. 
##  7 stl    2010 Aug                              0 -10.6   -175.    186.    175. 
##  8 stl    2010 Sep                              0  -7.52   300.   -293.   -300. 
##  9 stl    2010 Oct                              0  -4.45   258.   -254.   -258. 
## 10 stl    2010 Nov                              0  -1.39  -196.    197.    196. 
## # ... with 146 more rows, and abbreviated variable names
## #   1: `Amount of Crude Oil (Thousand Barrels)`, 2: season_year, 3: remainder,
## #   4: season_adjust
\end{verbatim}

\begin{verbatim}
## # A dable: 156 x 7 [1M]
## # Key:     .model [1]
## # :        Amount of total Petroleum Porducts (Thousand Barrels) = trend +
## #   season_year + remainder
##    .model     Date Amount of total Petroleum Po~1  trend seaso~2 remai~3 seaso~4
##    <chr>     <mth>                          <dbl>  <dbl>   <dbl>   <dbl>   <dbl>
##  1 stl    2010 Feb                           5360  8579.  -4551.   1332.   9911.
##  2 stl    2010 Mar                           6323  9169.  -2557.   -290.   8880.
##  3 stl    2010 Apr                           8564  9759.    954.  -2150.   7610.
##  4 stl    2010 May                          10587 10349.   -116.    354.  10703.
##  5 stl    2010 Jun                          11776 10911.    618.    247.  11158.
##  6 stl    2010 Jul                          11143 11472.   1216.  -1544.   9927.
##  7 stl    2010 Aug                          12516 12033.   1104.   -621.  11412.
##  8 stl    2010 Sep                          13405 12570.   1109.   -274.  12296.
##  9 stl    2010 Oct                          16466 13107.    822.   2537.  15644.
## 10 stl    2010 Nov                          18301 13644.    567.   4090.  17734.
## # ... with 146 more rows, and abbreviated variable names
## #   1: `Amount of total Petroleum Porducts (Thousand Barrels)`, 2: season_year,
## #   3: remainder, 4: season_adjust
\end{verbatim}

\begin{verbatim}
## # A dable: 156 x 7 [1M]
## # Key:     .model [1]
## # :        Amount of Natural gas (BOE) = trend + season_year + remainder
##    .model     Date `Amount of Natural gas (BOE)` trend season~1 remain~2 seaso~3
##    <chr>     <mth>                         <dbl> <dbl>    <dbl>    <dbl>   <dbl>
##  1 stl    2010 Feb                             0 38.2    -39.0     0.829   39.0 
##  2 stl    2010 Mar                             0 33.2    -39.4     6.19    39.4 
##  3 stl    2010 Apr                             0 28.2    -20.1    -8.17    20.1 
##  4 stl    2010 May                             0 23.2    -16.4    -6.87    16.4 
##  5 stl    2010 Jun                             0 19.2    -59.8    40.7     59.8 
##  6 stl    2010 Jul                             0 15.1    -40.8    25.7     40.8 
##  7 stl    2010 Aug                             0 11.0     -6.19   -4.78     6.19
##  8 stl    2010 Sep                             0  7.98   -14.1     6.07    14.1 
##  9 stl    2010 Oct                             0  4.99   -29.0    24.0     29.0 
## 10 stl    2010 Nov                             0  2.01   108.   -110.    -108.  
## # ... with 146 more rows, and abbreviated variable names 1: season_year,
## #   2: remainder, 3: season_adjust
\end{verbatim}

\includegraphics{draft1_files/figure-latex/decomposition_plots-1.pdf}
\includegraphics{draft1_files/figure-latex/decomposition_plots-2.pdf}
\includegraphics{draft1_files/figure-latex/decomposition_plots-3.pdf}

@@

\hypertarget{the-forecasters-toolbox}{%
\subsection{4. The forecaster's toolbox}\label{the-forecasters-toolbox}}

\hypertarget{exploring-different-simple-forecasting-methods}{%
\paragraph{4.1. Exploring different simple forecasting
methods}\label{exploring-different-simple-forecasting-methods}}

\hypertarget{exponential-smoothing}{%
\subsection{5. Exponential smoothing}\label{exponential-smoothing}}
